---
title: "Lab 1 output"
author: "Yanzhe Qu"
date: "2024-10-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solution:
Here we import our data:
```{r}

df = read.csv("/Users/alfonsoqu/Desktop/ASU/2024Fall/DAT 301/Lab 1/Absenteeism_at_work.csv", 
              sep=";", header = TRUE)

```

We use these commands: head and str, in order to have a basic understanding about out data set.
```{r}
head(df, 3)
str(df)
```

##Question 1
Plot the scatter plot of height vs. weight
```{r}
x = df$Weight
y = df$Height
plot(x,y,pch = 16, xlab="Weight", 
                   ylab="Height", 
                   main = "Height vs Weight")
```

It seems that as the weight increases the height increases as well.
```{r}
cor(x,y)
```

\break
##Question 2
Plot the histogram of hours of absences.
```{r}
Hours = df$Absenteeism.time.in.hours
hist(Hours, main = "Histogram of Absence Hours")
```

\break
We use more breaks for a clear visualization
```{r}
Hours = df$Absenteeism.time.in.hours
hist(Hours, main = "Histogram of Absence Hours", breaks = 120)
```

We can find out how manay people absent for at most 10 hours.
```{r}
sum(Hours <= 10)
length(Hours[Hours <= 10])
```

\break
##Question 3
Plot the histogram of age of persons of absences
```{r}
Age = df$Age
hist(Age, main = "Histogram of Age of Absence")
```

\break
##Question 4
Plot the bar of absence hours by month
```{r}
df_filtered <- df[df$Month.of.absence != 0, ]
sum <- tapply(X=df_filtered$Absenteeism.time.in.hours, INDEX=df_filtered$Month, FUN=sum)
barplot(sum, ylim=c(0,900), col="lightyellow", main="Frequency of Absent Hours by Month") 

```

\break
##Question 5
Box plot absence hours by social smoker
```{r}
df_filtered <- df[df$Absenteeism.time.in.hours != 0, ]

boxplot(df_filtered$Absenteeism.time.in.hours~df_filtered$Social.smoker,
        main = "Box Plot of Absence Hours by Social Smoker Status",
        xlab = "Social Smoker Absence Status",
        ylab = "Absence in Hours",
        col = c("lightblue", "salmon"),
        names = c("Non-Smokers", "Smokers"),
        las = 1,
        ylim = c(0, 30))
```
As shown in the plot, social smoker does not have a highter absence rate than non-smokers.

\break
##Question 6
Box plot absence hours by social drinker
```{r}
df_filtered <- df[df$Absenteeism.time.in.hours != 0, ]

boxplot(df_filtered$Absenteeism.time.in.hours~df_filtered$Social.drinker,
        main = "Box Plot of Absence Hours by Social Drinker Status",
        xlab = "Social Drinker Absence Status",
        ylab = "Absence in Hours",
        col = c("lightgreen", "orange"),
        names = c("Sobers", "Drinkers"),
        las = 1,
        ylim = c(0, max(df$Absenteeism.time.in.hours, na.rm = TRUE) + 10))
```
As shown in the plot, social drinker have a little bit more absence rate than sobers.